---
title: "What Makes New Games Stand Out on PC?"
subtitle: "An Inquiry on Steam's Top Selling New Games"
author: "Belal Yousofzai"
date: "2025-12-31"
date-format: "DD MMM YYYY"
title-block-banner: true
format: 
  html:
    lang: "en-GB"
    fig-format: svg
    toc: true
    toc-title: "Contents"
    theme:
      light: "default"
      dark: "darkly"
server: shiny
---


# Introduction

Thousands of exciting new video games are published for PCs every year. Among these 
new video games, their characteristics and features widely vary. Yet, only a small 
number of them sell well enough to be financially noteworthy. In particular, the 
most popular PC video game platform, Steam, showcases its best selling new video 
games on a yearly basis. In this article, we will examine Steam's best selling 
video games across different categories to understand what it is that potentially 
makes them sell so well.


## Why Focus on Steam?

Created by Valve Software in 2003, Steam has served as a convenient service to 
distribute and play PC video games.[^1] Since its inception, Steam has become the dominant 
video game platform on PC with more than 36 million registered users[^2] 
across the world[^3] and has amassed a library of more than 30 thousand games[^4].

While there are other online game platforms on PC, only a few are on a comparable 
scale to Steam in terms of library size and registered user count. The most significant 
ones are Microsoft's Xbox service on PC and Epic Game's Epic Game Store.

In the case of Microsoft, it has increasingly offered its once-exclusive Xbox games
on Steam in recent years, such that almost all of its games are readily available 
on Steam now. Third party games are usually released simultaneously on Steam and 
Xbox, so sales data will likely be identical across both platforms. More importantly 
for this analysis, the data on Xbox's best selling new games on PC is not easily 
accessible like Steam's data.

As for the Epic Game Store, the number of third party games being published on its 
platform is minuscule when compared to Steam. The Epic Game Store is significantly 
younger than Steam, meaning that the data available does not go so far back. 
Although the Epic Game Store has a larger PC user base of 295 million[^5], most 
of these users are concentrated in playing a few pre-existing games, one of which 
being Fortnite. In essence, the Epic Games Store lacks robust data for an analysis 
of the most popular new games sold on PC.

[^1]: About Valve https://www.valvesoftware.com/en/about
[^2]: Steam Daily User Count https://store.steampowered.com/charts
[^3]: Countries Steam is Used in https://store.steampowered.com/stats/content/
[^4]: About Steam https://store.steampowered.com/about/
[^5]: Epic Games Store 2024 Year In Review https://store.epicgames.com/en-US/news/epic-games-store-2024-year-in-review


# Finding the Best Selling Games

Every year, Steam announces its best games of the year in terms of revenue earned 
and peak player count. While Steam has been announcing its best games since 2017, 
only the webpages for 2022 and onward still exist. For the newly released games 
of a year, their ranking is calculated from the revenue earned in the first two 
weeks after launch[^6]. I assume Steam includes revenue earned from game keys sold 
by third party retailers as well. Now instead of specifying each game's revenue rank, 
the games are grouped into a top selling category of either platinum, gold, or silver.
From what I have seen, games that sold in the hundreds of thousands to low millions 
are listed as silver, games sold in the few millions or more are listed as gold, 
and games that sold beyond that are listed as platinum.

As of December 2025, Steam has added a bronze rank to its top selling categories.
This analysis will be updated in due time to include the recently released 2025 
data and additional bronze ranked games. Unfortunately, Steam has taken down its
webpage for 2021 and it is possible that the webpage for 2022 will eventually disappear.

Should you wish to see the best selling game lists directly, links are offered here:

[Best of 2025](https://store.steampowered.com/charts/bestofyear/2025)

[Best of 2024](https://store.steampowered.com/charts/bestofyear/2024)

[Best of 2023](https://store.steampowered.com/charts/bestofyear/2023)

[Best of 2022](https://store.steampowered.com/charts/bestofyear/2022)

[^6]: Explanation of Sales Ranking https://help.steampowered.com/en/faqs/view/6C17-2BC1-2A01-9B76


# Analysing the Games


## Acquiring the Data

To understand what makes the best selling games potentially stand out from the 
competition, their data needs be extracted for analysis. However, Steam does 
not seem to have official APIs for its store data. Thus, a web-scraping script was 
made in R to collect the data of all the best selling games. 2022 was selected as 
the starting year because 2021 lacked a silver category, which would have skewed 
the distribution of the data with extra gold and platinum games.

In making a web scraping script for this analysis, some unexpected challenges came 
about. The first was that the data points for each game were downloaded from their 
respective storefront webpage, which meant interacting with numerous web elements 
that were not always fixed in position. The next challenge was with handling absent 
values or values that did not follow a specified format. The last challenge was with 
Steam's age verification page acting as a blockade. After accounting for all these 
troubles, my script was able to successfully collect the data of the 150 games 
from the years 2022 to 2024. Further explanations of the web scraping issues 
encountered are covered in their respective sections.


## The Pricing

Let us begin with the prices of the games. Game prices were collected in USD 
because sales taxes are not included in the price and that games sold in developed 
markets have similar prices when exchange rates are factored in. The exchange 
rate table below is provided as a reference.

|Currency|Value|
|--------|-----|
|EUR     |0.85 |
|GBP     |0.75 |
|AUD     |1.55 |
|CAD     |1.40 |
|JPY     |155  |
|CNY     |7.10 |

: 1 USD is approximately equivalent to {.striped}

Now onto the graph. Free to play games and de-listed games are excluded from this 
section.

```{r}
#| label: fig-game_price
#| fig-cap: "Distribution of Game Price by Year"
#| out-width: "90%"
#| out-height: "80%"
#| warning: false

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

ggplot(data = filter(Steam_Data, price_USD != 0),
       aes(x = price_USD,
           fill = fct_relevel(factor(top_sell_category), 
                              "Platinum")
           ),
       width = "800px"
       ) +
  geom_histogram(binwidth = 10,
                 colour = rgb(0, 0, 0, 1)) +
  scale_fill_brewer() +
  scale_x_continuous(labels = scales::label_currency()) +
  facet_wrap(~year_published) +
  theme_bw() + 
  labs(title = "Game Price Distribution by Year",
       x = "USD Price",
       y = "Number of Games",
       fill = "Revenue Category")
```

From the above graph, we can see that even though game pricing is varied, a $60 
list price is the most common. Note that in 2024, numerous game companies did 
increase their standard retail price to $70. 
In the dataset, the median list price is $50 whereas the average list price is 
$48 ± 5 (depending on year and revenue category).


It is important to consider that games sold on Steam tend to have regional pricing 
in place. Regional pricing is to price a game in line with the economic conditions 
of a country, rather than at a price that was simply converted from the base
currency of the game. The effect of regional pricing is at its greatest for games 
sold in developing countries. Taking an example game from the dataset: 
[Black Myth Wukong](https://store.steampowered.com/app/2358720/) is listed at 
$60 USD. However, in its home market China the price is 268 CNY which is 37.75 USD, 
that is a 37% reduction.


## Outlier Games

Some of the games in the dataset are either free to own or unavailable to purchase.
Let us take a look at these games before moving on to the next sections.

### Free to Play Games

Since only 5% of the games from the dataset are free to play, they are presented 
in a table. The main highlight is that all the games are action oriented, with 
two of the games having card-based gameplay. Interestingly, the user ratings are 
not the best for these games. It's likely because players who otherwise would not 
play these games still give them a try because they are free. Or these free to play
games may not have the friendliest monetisation schemes.

```{r}
tableOutput("freeGamesTable")
```

```{r}
#| context: server

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

Steam_Data <- Steam_Data %>% mutate(
  user_ratings_total = user_ratings_positive + user_ratings_negative)

Steam_Data <- Steam_Data %>% mutate(
  user_ratings_pct = round(
    (user_ratings_positive / (user_ratings_positive + user_ratings_negative)), 
    digits = 2))

free_games_data <- Steam_Data %>% 
  filter(price_USD == 0) %>%
  arrange(year_published, month_published) %>%
  mutate(year_published = as.integer(year_published),
         age_rating_PEGI = as.integer(age_rating_PEGI)) %>%
  select(title, 
         publisher, 
         month_published, 
         year_published, 
         top_sell_category, 
         age_rating_PEGI, 
         user_ratings_pct) %>%
  rename(Title = title,
         Publisher = publisher,
         "Month Published" = month_published,
         "Year Published" = year_published,
         "Revenue Category" = top_sell_category,
         "Age Rating" = age_rating_PEGI,
         "User Rating %" = user_ratings_pct)


output$freeGamesTable <- renderTable(free_games_data, 
                                     striped = TRUE,
                                     hover = TRUE)

```


### Delisted Games

7% of the best selling games from the dataset have already been withdrawn from 
Steam's storefront. Except for 
[Multiversus](https://store.steampowered.com/app/1818750/), all the games are 
about sports or motorsports. As for why the games have been removed so soon, it 
could simply be because of licensing restrictions. Another reason could be that 
it is more profitable to have players buy a whole new game every year than to 
release add-ons for the existing games.

```{r}
tableOutput("delistedGamesTable")
```

```{r}
#| context: server

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

Steam_Data <- Steam_Data %>% mutate(
  user_ratings_total = user_ratings_positive + user_ratings_negative)

Steam_Data <- Steam_Data %>% mutate(
  user_ratings_pct = round(
    (user_ratings_positive / (user_ratings_positive + user_ratings_negative)), 
    digits = 2))

delisted_games_data <- Steam_Data %>% 
  filter(is.na(price_USD)) %>%
  arrange(year_published, month_published) %>%
  mutate(year_published = as.integer(year_published),
         age_rating_PEGI = as.integer(age_rating_PEGI)) %>%
  select(title, 
         publisher, 
         month_published, 
         year_published, 
         top_sell_category, 
         age_rating_PEGI, 
         user_ratings_pct) %>%
  rename(Title = title,
         Publisher = publisher,
         "Month Published" = month_published,
         "Year Published" = year_published,
         "Revenue Category" = top_sell_category,
         "Age Rating" = age_rating_PEGI,
         "User Rating %" = user_ratings_pct)

output$delistedGamesTable <- renderTable(delisted_games_data, 
                                         striped = TRUE,
                                         hover = TRUE)

```


## The Genres

On Steam, publishers give their games a set of official genre tags to help players 
know more about the games they intend to play. Below is an interactive graph that 
lets you view the most common genre tags among the games in the dataset.


```{r}
#| panel: sidebar

selectInput("genreCategory", 
            label = "Select Category", 
            choices = list(
              "All",
              "Platinum",
              "Gold",
              "Silver"
            ),
            selected = "All"
)

selectInput("genreYear", 
            label = "Select Year", 
            choices = list(
              "All",
              2022,
              2023,
              2024
            ),
            selected = "All"
)

plotOutput("genrePlot")

```

```{r}
#| panel: fill
plotOutput("genrePlot",
           height = "600px")
```

```{r}
#| context: server

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

output$genrePlot <- renderPlot(
  {
    games <- Steam_Data
    plot_subtitle <- "For All Games"
    if (!str_detect(input$genreYear, "All"))
    {
      games <- games %>% filter(year_published == as.numeric(input$genreYear))
      plot_subtitle <- paste0("For ", input$genreYear)
    }
    if (!str_detect(input$genreCategory, "All"))
    {
      games <- games %>% filter(top_sell_category == input$genreCategory)
      plot_subtitle <- paste0("For ", input$genreCategory)
    }
    # Put all individual tags of the games into a table
    all_genre_tags <- c()
    for (raw_genre_tags in games$genre)
    {
      tags <- str_split_1(raw_genre_tags, coll("|"))
      all_genre_tags <- append(all_genre_tags, tags) 
    }
    genre_tags_data <- sort(table(all_genre_tags), decreasing = FALSE)
    # Plot the data
    ggplot(data = data.frame(genre_tags_data),
           aes(y = all_genre_tags,
               x = Freq)) +
      geom_col(width = 0.7,
               position = "dodge2",
               fill = rgb(0.15, 0.26, 0.65),
               show.legend = FALSE) +
      theme_classic() +
      theme(axis.title.y = element_blank()) +
      labs(title = "Most Popular Genre Tags",
           subtitle = plot_subtitle,
           y = "Genre",
           x = "Number of Games in the Genre")
  },
  res = 95
)

```

From the above interactive graph, the most popular genre tags by far are action, 
adventure, and RPG. Action essentially means a game with challenges, such as combat. 
Adventure refers to a game that has a story. RPG stands for Role Playing Game. 
RPGs can be thought of as adventure games that offer deep character customisation.

Why would action games be the most popular? For a start, action as a subject matter
can be conveyed well in the form a a video game. Whereas action in films 
can be rough or short lived, and limited in books, video games allow players the 
opportunity to engage in continuous, safe, and fun action. Furthermore, if we assume
that the general demographic of Steam consists of young men, then having the action 
genre be the most popular is self explanatory.

To give some examples of what games from the different genres are like, a representative 
action game can be [Tekken 8](https://store.steampowered.com/app/1778820/). 
While the game somewhat has a story, it is primarily about two opponents engaged 
in intense hand-to-hand combat. A representative adventure game can be 
[Stray](https://store.steampowered.com/app/1332010/). The game is about a 
lonesome cat uncovering the mysteries of a grotty futuristic city. A representative 
RPG game can be [Dragon Quest III](https://store.steampowered.com/app/2701660/). 
That game is about a young hero and his (or her) allies who set off on a 
grand quest to defeat the villainous fiend that threatens their fantasy world.
Genre tags can be treated as suggestions on what a game is like. Even Steam does 
not seem to have strong guidelines on what genres a game should be put into.


## The Publishers

92 publishers released the 150 games in the dataset. However, only some of the 
publishers had more than one best selling game. Rather than overwhelming ourselves
with these one-off publishers, let us examine the more successful publishers in 
the interactive graph below. Only 10 publishers are shown, sorted by the number 
of games they have in the dataset. 

```{r}
#| panel: sidebar

selectInput("publisherCategory", 
            label = "Select Category", 
            choices = list(
              "All",
              "Platinum",
              "Gold",
              "Silver"
            ),
            selected = "All"
)

selectInput("publisherYear", 
            label = "Select Year", 
            choices = list(
              "All",
              2022,
              2023,
              2024
            ),
            selected = "All"
)

plotOutput("publisherPlot")

```

```{r}
#| panel: fill
plotOutput("publisherPlot",
           height = "600px")
```

```{r}
#| context: server
#| panel: fill

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

output$publisherPlot <- renderPlot(
  {
    max_publishers <- 12 # There are too many publishers to show, limit it
    min_games <- 0 # Many publishers only appear once, exclude them
    publisher_data <- Steam_Data
    if (!str_detect(input$publisherYear, "All"))
      publisher_data <- publisher_data %>% 
      filter(year_published == input$publisherYear)
    if (!str_detect(input$publisherCategory, "All"))
      publisher_data <- publisher_data %>% 
      filter(top_sell_category == input$publisherCategory)
    publisher_data <- publisher_data %>%
      count(publisher) %>%
      filter(n > min_games) %>%
      arrange(n) %>%
      slice_tail(n = max_publishers)
    publisher_tibble <- as_tibble(publisher_data)
    max_publish_count <- max(publisher_tibble$n)
    # Plot the data
    ggplot(data = publisher_data,
           aes(x = reorder(publisher, n),
               y = n)) +
      geom_col(fill = rgb(0.15, 0.26, 0.65)) +
      geom_text(aes(label = paste0(n), hjust = -0.5)) +
      scale_y_continuous(limits = c(0, max_publish_count * 1.1)) +
      coord_flip() +
      theme_classic() +
      theme(axis.title.y = element_blank()) +
      labs(title = "Most Popular Publishers",
           x = "Publisher",
           y = "Number of Games Listed")
  },
  res = 95
)
```

The most popular publisher overall is Electronic Arts (EA). This is mainly because 
EA owns some of the most globally popular video game series. To name a few of the 
series: FIFA, Formula 1, The Sims, Battlefield, and Need for Speed. Considering 
that EA already has its own PC platform, Origin/EA App, the additional strong sales 
of its games on Steam signifies its dominance in the PC game market.

PlayStation and SEGA follow EA in having the most best selling games in the dataset.
PlayStation's appearance is significant. Until recently, games published by PlayStation 
were only available on its own consoles. I believe that Sony Interactive Entertainment, 
the company in charge of PlayStation, recognised that the PC gaming market has 
matured to the stage where convincing a sizeable portion of current PC players to 
get a PlayStation 5 to play exclusive games would not be economically possible. 
In a similar move to Microsoft, Sony has released most of its commercially successful 
games onto Steam over the past 5 years.

Sony and Microsoft are not alone in bringing over their exclusive games to Steam. 
SEGA, Square Enix, Ubisoft, and other large publishers have been releasing more 
of their existing games on Steam. In recent years, many of these publishers now 
simultaneously launch their games on consoles and Steam. Previously, particularly 
before the Covid-19 pandemic, many large publishers would only release some 
of their games on PC one year after their release on consoles and they would often 
be available only on their own distribution platforms. Even though PC game platform 
competition is now dwindling, online sentiment of Steam's natural monopolisation 
has been positive. PC players generally prefer Steam when given the choice.


## The Age Ratings

Now for the age ratings. In the graph below, games rated for teenagers and adults 
sold the most on Steam. This is understandable as Steam is meant for users aged 13 
and above. Most of the age 3 games in the dataset are just sports games. Games 
primarily meant for children do not seem to be that popular on Steam. If we were 
to think about the popular games children usually play, they are often played on 
mobile devices, game consoles such as the Nintendo Switch or PlayStation 4/5, or 
on PC platforms other than Steam.

```{r}
#| label: fig-game_age_ratings
#| fig-cap: "PEGI Age Ratings by Year and Revenue Category"
#| out-width: "90%"
#| out-height: "80%"
#| warning: false

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

ggplot(data = filter(Steam_Data, !is.na(age_rating_PEGI)),
       aes(x = factor(age_rating_PEGI),
           fill = fct_relevel(factor(top_sell_category), 
                              "Platinum"))) +
  geom_bar(colour = rgb(0, 0, 0)) +
  scale_fill_brewer() +
  facet_wrap(~year_published) +
  theme_bw() + 
  labs(title = "Game Age Rating Distribution",
       subtitle = "By Year and Revenue Category",
       x = "PEGI Age Rating",
       y = "Number of Games",
       fill = "Revenue Category")

```

To learn more about PEGI, click [here](https://pegi.info). The PEGI age rating system
is made for Europe. However, it is used in some other regions of the world as well. 
A PEGI rating recommends the minimum age for a video game based on how appropriate 
its subject matter is. PEGI was chosen because of its intuitiveness and popularity. 
PEGI's age ratings tend to be in line with other organisations' ratings. To learn 
more about the age rating systems of other organisations, click [here](https://www.globalratings.com).

When collecting the PEGI age ratings for the games on Steam, I discovered that many
of the games surprisingly did not have an age rating label on Steam. Thus, I had 
to manually fill in the missing values with the ones from PEGI's website. Still, 
not all the games in the dataset have been rated by PEGI, so some missing values 
are to be expected.


## The Release Dates

Let us find out what are the most and least popular months to publish a game on 
Steam in the interactive graph below.

```{r}
#| panel: sidebar

selectInput("releaseCategory", 
            label = "Select Category", 
            choices = list(
              "All",
              "Platinum",
              "Gold",
              "Silver"
            ),
            selected = "All"
)

selectInput("releaseYear", 
            label = "Select Year", 
            choices = list(
              "All",
              2022,
              2023,
              2024
            ),
            selected = "All"
)

plotOutput("releasePlot")

```

```{r}
#| panel: fill
plotOutput("releasePlot",
           height = "600px")
```

```{r}
#| context: server

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

output$releasePlot <- renderPlot(
  {
    release_data <- Steam_Data
    if (!str_detect(input$releaseYear, "All"))
      release_data <- release_data %>% 
        filter(year_published == input$releaseYear)
    if (!str_detect(input$releaseCategory, "All"))
      release_data <- release_data %>% 
        filter(top_sell_category == input$releaseCategory)
    ggplot(data = release_data, 
           aes(x = month_published)) +
      geom_bar(fill = rgb(0.15, 0.26, 0.65)) +
      theme_classic() +
      labs(x = "Month",
           y = "Number of Games Released")
  },
  res = 95
)

```

Surprisingly, September is the most popular month overall. Considering 
that is the month many people return to school in, I was expecting the summer 
or the end of the year to be the most popular. Interestingly, July is the least popular 
month for publishing games. I do not have a definite answer for the results here, 
so this is something worth following up on.


## The Languages

On Steam, the language support of a game is recognised by three categories: interface,
subtitles, and full audio. The interface consists of the text elements of a game, 
such as the menus, manuals, and safety warnings. Subtitles are the descriptive texts 
of a game's audio. Full audio means that the game has voiced dialogue.

In the dataset, I found that if a game has a supported interface language, then 
the subtitles (if they exist) will be available in that supported language. Thus, 
support for interfaces and subtitles overlap. Not all games in the dataset have 
any voiced dialogue, so their support for full audio is blank. Let us examine the 
supported languages further in the interactive graph below.

```{r}
#| panel: sidebar

selectInput("languageType", 
            label = "Select Audio Type", 
            choices = list(
              "Interface/Subtitles",
              "Full Audio"
            ),
            selected = "Interface/Subtitles"
)

selectInput("languageCategory", 
            label = "Select Category", 
            choices = list(
              "All",
              "Platinum",
              "Gold",
              "Silver"
            ),
            selected = "All"
)

selectInput("languageYear", 
            label = "Select Year", 
            choices = list(
              "All",
              2022,
              2023,
              2024
            ),
            selected = "All"
)

plotOutput("languagesPlot")

```

```{r}
#| panel: fill
plotOutput("languagesPlot",
           height = "610px")
```

```{r}
#| context: server

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

# Get the language frequency of the games in the dataset
get_lang_freq <- function(steam_data, lang_type, year, revenue_category)
{
  if (!str_detect(year, "All")) 
    steam_data <- steam_data %>% filter(year_published == year)
  if (!str_detect(revenue_category, "All")) 
    steam_data <- steam_data %>% filter(top_sell_category == revenue_category)
  all_langs <- c()
  if (str_detect(lang_type, "Interface/Subtitles")) 
    choice <- steam_data$languages_interface
  else if (str_detect(lang_type, "Full Audio")) 
    choice <- steam_data$languages_spoken
  for (raw_langs in choice)
  {
    if (is.na(raw_langs)) next
    langs <- str_split_1(raw_langs, coll("|"))
    all_langs <- append(all_langs, langs) 
  }
  return(sort(table(all_langs), decreasing = FALSE))
}

output$languagesPlot <- renderPlot(
  {
    filtered_data <- Steam_Data
    if (str_detect(input$languageType, "Interface/Subtitles"))
    {
      filtered_data <- filter(filtered_data, !is.na(Steam_Data$languages_interface))
      game_count <- as.numeric(count(filtered_data))
    }
    if (str_detect(input$languageType, "Full Audio"))
    {
      filtered_data <- filter(filtered_data, !is.na(Steam_Data$languages_spoken))
      game_count <- as.numeric(count(filtered_data))
    }
    if (!str_detect(input$languageCategory, "All"))
    {
      filtered_data <- filter(filter(filtered_data, 
                                     filtered_data$top_sell_category == input$languageCategory))
      game_count <- as.numeric(count(filtered_data))
    }
    if (!str_detect(input$languageYear, "All"))
    {
      filtered_data <- filter(filtered_data, 
                              filtered_data$year_published == input$languageYear)
      game_count <- as.numeric(count(filtered_data))
    }
    language_data <- get_lang_freq(Steam_Data, 
                                   input$languageType, 
                                   input$languageYear, 
                                   input$languageCategory)
    # Plot the data
    ggplot(data = data.frame(language_data),
           aes(y = all_langs,
               x = Freq)) +
      geom_col(width = 0.7,
               position = "dodge2",
               fill = rgb(0.15, 0.26, 0.65),
               show.legend = FALSE) +
      geom_text(aes(label = paste0(round(Freq/game_count * 100), "%"), 
                    hjust = -0.2)
      ) +
      scale_x_continuous(limits = c(0, max(language_data) * 1.1)) +
      theme_classic() +
      theme(axis.title.y = element_blank()) +
      labs(title = "Language Support in Games",
           #y = "Languages",
           x = "Number of Games Supporting the Language")
  },
  res = 95,
  execOnResize = TRUE
)

```

While all the games in the dataset support English interfaces, almost all of them 
support French and German interfaces. It is a pleasure to see so many games be available 
in multiple languages. Perhaps in this case it is the least common languages 
that are the most interesting ones. [Battlebit Remastered](https://store.steampowered.com/app/671860/) 
is the one game in the dataset that supports Belarusian, Estonian, Macedonian, 
and Slovak. Only [Farming Simulator 25](https://store.steampowered.com/app/2300320/) 
supports Romanian. Three Playstation games support Croatian, specifically 
[The Last of Us Part I](https://store.steampowered.com/app/1888930/), 
[Ghost of Tschushima Directors Cut](https://store.steampowered.com/app/2215430/), 
and [God of War Ragnarok](https://store.steampowered.com/app/2322010/).
For comparison, Steam's interface is available in 26 languages.[^7]

For Full Audio, only 
[Microsoft Flight Simulator 2024](https://store.steampowered.com/app/2537590/) 
(MSFS 24) supports Finnish, Norwegian, and Swedish. I believe that for MSFS 24, 
the majority of its voices are produced by AI, that was the case for MSFS 2020 at 
least. [S.T.A.L.K.E.R. 2: Heart of Chornobyl](https://store.steampowered.com/app/1643320/) 
is the one game with Ukrainian full audio support. The Last of Us Part I and God 
of War Ragnarok have Greek full audio support. The Last of Us Part I and MSFS 24 
are the only games to have Turkish full audio support.

[^7]: Steam Language-Support Information https://partner.steamgames.com/doc/store/localization/languages


## The Recommended Graphics Cards

For every game on Steam, publishers or developers can list the minimum and recommended
graphics cards for their games. Usually, if a player's PC meets the recommended 
requirements, then he or she will be able to enjoy the game on high graphics setting 
at a resolution of 1920x1080 at 60 frames per second. The interactive graph below 
reveals the most commonly recommended graphics cards for the games in the dataset.

```{r}
#| panel: sidebar

selectInput("graphicsCompany", 
            label = "Select Company", 
            choices = list(
              "All",
              "AMD",
              "Intel",
              "Nvidia"
            ),
            selected = "All"
)

selectInput("graphicsCategory", 
            label = "Select Category", 
            choices = list(
              "All",
              "Platinum",
              "Gold",
              "Silver"
            ),
            selected = "All"
)

selectInput("graphicsYear", 
            label = "Select Year", 
            choices = list(
              "All",
              2022,
              2023,
              2024
            ),
            selected = "All"
)

plotOutput("graphicsPlot")

```

```{r}
#| panel: fill
plotOutput("graphicsPlot",
           height = "600px")
```

```{r}
#| context: server

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

# Get the recommended GPUs frequency from the dataset
get_gpu_freq <- function(steam_data, year, revenue_category)
{
  if (!str_detect(year, "All")) 
    steam_data <- steam_data %>% filter(year_published == year)
  if (!str_detect(revenue_category, "All")) 
    steam_data <- steam_data %>% filter(top_sell_category == revenue_category)
  all_gpus <- c()
  for (raw_gpus in steam_data$recommended_graphics)
  {
    if (is.na(raw_gpus)) next
    gpus <- str_split_1(str_to_upper(raw_gpus), coll("|"))
    all_gpus <- append(all_gpus, gpus) 
  }
  return(as_tibble(table(all_gpus)))
}

# Filter GPUs by company
get_gpus_by_company <- function(gpu_data, gpu_company)
{
  max_gpus_to_show <- 12 # Many GPUs only appear once, exclude them for brevity
  name_start <- 1L
  gpu_company <- str_to_upper(gpu_company)
  if (!str_detect(gpu_company, "ALL"))
  {
    gpu_data <- gpu_data %>% 
      filter(str_detect(all_gpus, gpu_company))
  }
  if (str_detect(gpu_company, "NVIDIA")) name_start <- 16L
  else if (str_detect(gpu_company, "AMD")) name_start <- 12L
  else if (str_detect(gpu_company, "INTEL")) name_start <- 7L
  gpu_data <- gpu_data %>% 
    arrange(n) %>%
    mutate(all_gpus = str_sub(all_gpus, start = name_start)) %>%
    slice_tail(n = max_gpus_to_show) %>%
    rename(gpu_name = all_gpus)
  return(gpu_data)
}

output$graphicsPlot <- renderPlot(
  {
    graphics_data <- get_gpu_freq(Steam_Data, 
                                  input$graphicsYear, 
                                  input$graphicsCategory)
    graphics_data <- get_gpus_by_company(graphics_data, input$graphicsCompany)
    ggplot(data = graphics_data,
           aes(y = fct_reorder(gpu_name, n),
               x = n)) +
      geom_col(fill = rgb(0.15, 0.26, 0.65),
               show.legend = FALSE) +
      theme_classic() +
      theme(axis.title.y = element_blank()) +
      labs(title = "Most Recommended Graphics Cards",
           y = "Graphics Card",
           x = "Number of Games Recommending the Graphics Card")
  },
  res = 95
)

```

To summarise the findings, graphic cards that are designed to play the latest games 
(at the time of their release) at 1920x1080 or 2560x1440 on high graphics settings 
are the most commonly recommended graphics cards. Currently, the most commonly 
owned graphics card among Steam users is the Nvidia GeForce RTX 4060 for laptops.[^8] 
That is a formidable processor. Therefore, many Steam players will be able to enjoy 
the majority of the best selling games in the dataset at the recommended graphics 
settings.

One issue with collecting this data is that publishers are not obligated to 
follow a standard format when writing out the requirements for their games. Thus,
a lot of parsing and manual correcting were needed to make the graphics cards data 
uniform. As an example, some game requirements would have the "Nvidia GeForce 
GTX 1660" abbreviated as "GTX 1660".

[^8]: Hardware Used by Steam Users: https://store.steampowered.com/hwsurvey/


## The User Ratings

On Steam, users can leave a review that either recommends or does not recommend 
the game they played. From the dataset, more than 5 million reviews posted within 
one month of a game's release date were counted with 4.1 million of them being 
positive (81.85%) and 900 thousand being negative (18.15%).

To read the box plot figure below, the coloured boxes represent the middle 50% 
of the user rating data points. The black line in the coloured box is the median 
of the data points. The lines extending out of the boxes represent the minimum and 
maximum values of the data points. Data points placed beyond the lines are 
outliers. A stretched out box with long lines implies that the data points 
are quite scattered. For a greater explanation of box plots,
click [here](https://wellbeingatschool.org.nz/information-sheet/understanding-and-interpreting-box-plots).

```{r}
#| label: fig-game_user_ratings
#| fig-cap: "Steam User Ratings by Year and Revenue Category"
#| out-width: "90%"
#| out-height: "80%"
#| warning: false

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

# Calculate User Rating, add it as a dedicated column to the dataset
Steam_Data <- Steam_Data %>% mutate(
  user_ratings_total = user_ratings_positive + user_ratings_negative)

Steam_Data <- Steam_Data %>% mutate(
  user_ratings_pct = round(
    (user_ratings_positive / (user_ratings_positive + user_ratings_negative)), 
    digits = 2))

ggplot(data = Steam_Data,
       aes(x = as.character(year_published),
           y = user_ratings_pct,
           fill = fct_relevel(factor(top_sell_category), 
                              "Platinum"))) +
  geom_boxplot(width = 0.8) +
  annotate("text", x = 1.98, y = 0.13, label = "NBA 2K24") +
  annotate("text", x = 2.95, y = 0.38, label = "MSFS 2024") +
  scale_fill_brewer() +
  scale_y_continuous(breaks = c(0.2, 0.4, 0.6, 0.8, 1),
                     guide = guide_axis(minor.ticks = TRUE),
                     labels = scales::label_percent()) +
  theme_classic() +
  labs(title = "User Ratings",
       subtitle = "By Year and Revenue Category",
       x = "Year",
       y = "Recommendation",
       fill = "Revenue Category")

```

Overall, most games tend to have a rating in the range of 60% to 90%.
Regarding the outliers in the above figure, we have two of them: 
[NBA 2K24](https://store.steampowered.com/app/2338770/) at 13% 
and Microsoft Flight Simulator 2024 (MSFS 24) at 38%. NBA 2K24 is by far the worst 
rated game in the dataset and is one of the lowest rated games on Steam to date. 
MSFS 24, while low, ended up as an outlier because the rest of the games in its 
category were rated significantly higher. Generally, PC players post negative 
ratings when they experience technical problems such as audio glitches or visual 
stuttering. For MSFS 24, many of its user reviews complained about download and 
streaming problems. Many of the reviews also complained about the simulator feeling 
like an unfinished product.

To know more about the worst rated games in the dataset, below is a table of the 
games with a user rating less than 60%. These games make up 14% of the dataset. 

```{r}
div(style="height:450px; overflow-y: scroll",
    tableOutput("userRatingTable")
)
```

```{r}
#| context: server

library(tidyverse)

Steam_Data <- read_csv("./Data/Steam_Top_Selling_New_Games.csv")

Steam_Data <- Steam_Data %>% mutate(
  user_ratings_total = user_ratings_positive + user_ratings_negative)

Steam_Data <- Steam_Data %>% mutate(
  user_ratings_pct = round(
    (user_ratings_positive / (user_ratings_positive + user_ratings_negative)), 
    digits = 2))

ratings_data <- Steam_Data %>% filter(user_ratings_pct < 0.6)

ratings_data <- ratings_data %>% arrange(user_ratings_pct)

ratings_data <- ratings_data %>% select(title, 
                                        publisher, 
                                        top_sell_category,
                                        user_ratings_pct,
                                        user_ratings_total,
                                        user_ratings_positive,
                                        user_ratings_negative
)

ratings_data <- ratings_data %>% mutate(user_ratings_total = 
                                          as.integer(user_ratings_total),
                                        user_ratings_positive =
                                          as.integer(user_ratings_positive),
                                        user_ratings_negative = 
                                          as.integer(user_ratings_negative)
)

ratings_data <- ratings_data %>% rename(Title = title,
                                        Publisher = publisher,
                                        "Revenue Category" = top_sell_category,
                                        "User Ratings %" = user_ratings_pct,
                                        "Total User Ratings" = user_ratings_total,
                                        "Positive User Ratings" = user_ratings_positive,
                                        "Negative User Ratings" = user_ratings_negative
)

output$userRatingTable <- renderTable(ratings_data, 
                                      striped = TRUE,
                                      hover = TRUE,
                                      spacing = "m")

```


# Conclusion

After examining the 150 best selling games on Steam, some noticeable features stand 
out. For a start, $60 USD games are most commonly sought after by Steam users. However, 
games across the price range are still being bought up. The action, adventure, 
and RPG genres are the most popular genres. Players enjoy a story with challenge 
and customisation. Large high-budget publishers produce many of the best selling 
games. Naturally, their games are designed for and marketed to a varied global 
audience. Games rated for ages 12 and up are purchased the most by Steam users. 
People looking to publish games designed for children may not expect to earn the 
greatest revenue on Steam.

The most popular month to publish a game in is September. Whereas the least popular 
month to publish a game in is July. Early Autumn, late Winter, and late Spring are 
the most common time periods for a game to be released. In terms of language support, 
the most popular languages are English, German, French, Spanish, Japanese, Chinese, 
and Korean. For games with voiced audio, English is the #1 most supported language.
The most commonly recommended graphics cards are approximately in line with the 
graphics cards owned by Steam users. Lastly, Steam users generally leave a positive 
recommendation for the best selling games they played.

If you are interested in exploring the dataset, click here to access it from Github.
